---
title: "Titanic"
author: "Nicole Wittlin, Cho Kim, Jonathan Marin"
date: "April 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Kaggle Titanic Competition

https://www.kaggle.com/c/titanic/data


# Introduction


The sinking of the RMS Titanic is one of the most infamous shipwrecks in history.  On April 15, 1912, during her maiden voyage, the Titanic sank after colliding with an iceberg, killing 1502 out of 2224 passengers and crew. This sensational tragedy shocked the international community and led to better safety regulations for ships.

One of the reasons that the shipwreck led to such loss of life was that there were not enough lifeboats for the passengers and crew. Although there was some element of luck involved in surviving the sinking, some groups of people were more likely to survive than others, such as women, children, and the upper-class.

In this kaggle challenge, we must complete the analysis of what sorts of people were likely to survive. In particular, we need to apply the tools of machine learning to predict which passengers survived the tragedy.

# Packages

```{r echo = TRUE, include = TRUE, warning = FALSE}

#library(xlsx)
library(knitr)
library(ggplot2)
library(sqldf)
library(reshape2)
library(magrittr)
library(kableExtra)
library(doBy)
library(caret)
library(parallel)
library(doParallel)
library(randomForest)

```

#Data Dictionary

```{r}

#variables <- read.xlsx("Variables.xlsx", "Sheet1")
variables <- read.csv("Variables.csv", sep=",")


kable((variables), "html") %>% kable_styling(bootstrap_options="striped", full_width = F, font_size = 12)

```


#Variable Notes

Variable Notes
pclass: A proxy for socio-economic status (SES)
1st = Upper
2nd = Middle
3rd = Lower

age: Age is fractional if less than 1. If the age is estimated, is it in the form of xx.5

sibsp: The dataset defines family relations in this way...
Sibling = brother, sister, stepbrother, stepsister
Spouse = husband, wife (mistresses and fiancés were ignored)

parch: The dataset defines family relations in this way...
Parent = mother, father
Child = daughter, son, stepdaughter, stepson
Some children travelled only with a nanny, therefore parch=0 for them.

#Import the Data

```{r echo = TRUE, include = TRUE}


##setwd("C:/Users/i64756/Desktop/Kaggle/Titanic")


train <- read.csv("train.csv")

#Adding a column to train to indicate that it is the train set

train$Set <- "train"

test <- read.csv("test.csv")

test$Set <- "test"

#We have two train and test sets.  One to try imputaing age, and the other to get it from the web. 
#We would like to see the difference. 

# age added for data with blanks from encylopedia-titanica.
train_age <- read.csv("trainage.csv", sep=",") 
train_age$Set <- "train"

test_age <- read.csv("testage.csv", sep = ",")
test_age$Set <- "test"

```


#Exploratory Analysis

## Women and Children First


```{r echo = TRUE, include = TRUE}

ggplot(train, aes(x=Age, group = Survived, fill = Survived)) +
    geom_histogram(binwidth=5, alpha=.5, position="identity")+   facet_grid(~Survived, labeller = label_both) + theme(legend.position="none")


ggplot(train, aes(x=Sex, group = Survived, fill = Survived)) +
    geom_histogram(stat = "count", binwidth=5, alpha=.5, position="identity")+   facet_grid(~Survived, labeller = label_both) + theme(legend.position="none")


ggplot(train, aes(x=Age, group = Survived, fill = Survived)) +
    geom_histogram(aes(y=(..count..)/sum(..count..)),binwidth=5, alpha=.5, position="identity")+   facet_grid(~Survived, labeller = label_both) + theme(legend.position="none") +
  ylab("Percent")


ggplot(train, aes(x=Sex, group = Survived, fill = Survived)) +
    geom_histogram(stat = "count", aes(y=(..count..)/sum(..count..)),binwidth=5, alpha=.5, position="identity")+   facet_grid(~Survived, labeller = label_both) + theme(legend.position="none") +
  ylab("Percent")


ggplot(train, aes(x=Sex, group = Survived, fill = Survived)) +
    geom_histogram(stat = "count", aes(y=(..count..)/sum(..count..)),binwidth=5, alpha=.5, position="identity")+   facet_grid(~Survived, labeller = label_both) + theme(legend.position="none") + ylab("Percent")
  

 




```

#Clean and Prep the Data

```{r}
#We want to add the Survived variable to the test set so that we can combine training and test 

test$Survived <- NA

test_age$Survived <- NA


#Combining the train and test set so that we can clean the data for both sets 
combined <- rbind(train,test)

combined_age <- rbind(train_age,test_age)

#Testing to see what class these features are

class(combined$PassengerId)
class(combined$Survived)
class(combined$Pclass)
class(combined$Name)
class(combined$Sex)
class(combined$Age)
class(combined$SibSp)
class(combined$Parch)
class(combined$Ticket)
class(combined$Fare)
class(combined$Embarked)

#I don't see any issues with the classes for the features

```

#Missing Variable Values

We need to see what values are missing and what to do with them.  With little data, we shouldn't exclude if we can.  Possibly we can impute them. 

```{r}

sapply(combined, function(x) sum(is.na(x)))

#The 418 surived variables make sense because we have the test set here.  However, the age is missing quite a bit.  Probably need to come up with a way to fill in the age by imputing. 



# Found a source that had ages 
sapply(combined_age, function(x) sum(is.na(x)))



```

#Feature Engineering

```{r}

#Want to create a variable by title.  This might help predict age. 

combined$Title <- gsub( "^.*, ", '' , combined$Name)
combined$Title <- gsub(" .*" , '', combined$Title)

combined_age$Title <- gsub( "^.*, ", '' , combined$Name)
combined_age$Title <- gsub(" .*" , '', combined$Title)


```

#Imputing Age

There are many ways to impute age and this is obviously an important variable to the model for two possible reasons:

1) We know that Women and Children were allowed on life boats first
2) Age is a factor in determing a person's health once on a boat which is crucial for survival. 

That being said, would it make sense to group age into smaller categories because being 22 vs being 23 might not be significant.  However, being an infant vs toddler vs teenager vs senior citizen might be. 

First, let's make a new variable for age to categorize.


```{r}


FilledAges <- sqldf("SELECT * FROM combined where Age is not null")

MissingAgesTitle <- sqldf("SELECT Title, count(*) as 'Count' FROM combined where Age is null group by Title")

MissingAgesTitle

summaryBy(Age ~ Title, data = FilledAges, FUN =  list(mean, max,min, median, sd))

#Imputed Age: Using the medians of Age per title

combined <- sqldf("select a.*,  case 
  when Age is null and Title = 'Capt.' then 70 
  when Age is null and Title = 'Col.' then 54 
  when Age is null and Title = 'Don.' then 40 
  when Age is null and Title = 'Dona.' then 39 
  when Age is null and Title = 'Dr.' then 49 
  when Age is null and Title = 'Lady' then 48
  when Age is null and Title = 'Major.' then 48
  when Age is null and Title = 'Master.' then 4
  when Age is null and Title = 'Miss.' then 22
  when Age is null and Title = 'Mlle.' then 24
  when Age is null and Title = 'Mme.' then 24
  when Age is null and Title = 'Mr.' then 29
  when Age is null and Title = 'Mrs.' then 40
  when Age is null and Title = 'Ms.' then 28
  when Age is null and Title = 'Rev.' then 42
  when Age is null and Title = 'Sir.' then 49
  else Age end as 'ImputedAge' from combined a")


combined_age <- sqldf("select a.*,  case 
  when Age is null and Title = 'Capt.' then 70 
  when Age is null and Title = 'Col.' then 54 
  when Age is null and Title = 'Don.' then 40 
  when Age is null and Title = 'Dona.' then 39 
  when Age is null and Title = 'Dr.' then 49 
  when Age is null and Title = 'Lady' then 48
  when Age is null and Title = 'Major.' then 48
  when Age is null and Title = 'Master.' then 4
  when Age is null and Title = 'Miss.' then 22
  when Age is null and Title = 'Mlle.' then 24
  when Age is null and Title = 'Mme.' then 24
  when Age is null and Title = 'Mr.' then 29
  when Age is null and Title = 'Mrs.' then 40
  when Age is null and Title = 'Ms.' then 28
  when Age is null and Title = 'Rev.' then 42
  when Age is null and Title = 'Sir.' then 49
  else Age end as 'ImputedAge' from combined_age a")

#Check to make sure all ages were imputed
sapply(combined, function(x) sum(is.na(x)))

sapply(combined_age, function(x) sum(is.na(x)))


```


#Creating Age Classes



```{r warning = FALSE}

combined <- sqldf("SELECT a.*, CASE WHEN ImputedAge is Null then NULL when ImputedAge <= 2 then 'Infant' when ImputedAge <= 5 then 'Toddler' when ImputedAge <= 12 then 'Child' when ImputedAge <= 18 then 'Teen' when ImputedAge <= 30 then 'Adult' when ImputedAge <= 40 then 'OlderAdult' when ImputedAge <= 65 then 'OvertheHill' else 'Senior' end as 'AgeClass' from combined a")

combined_age <- sqldf("SELECT a.*, CASE WHEN ImputedAge is Null then NULL when ImputedAge <= 2 then 'Infant' when ImputedAge <= 5 then 'Toddler' when ImputedAge <= 12 then 'Child' when ImputedAge <= 18 then 'Teen' when ImputedAge <= 30 then 'Adult' when ImputedAge <= 40 then 'OlderAdult' when ImputedAge <= 65 then 'OvertheHill' else 'Senior' end as 'AgeClass' from combined_age a")




#Check to make sure that I categorized all passengers
sapply(combined, function(x) sum(is.na(x)))


```

#Combine Rare Titles

After running a model with the titles created as is, there was an issue where the test set had a title that the train set did not. Will need to combine the rare titles. 

```{r warning = FALSE}

sqldf("Select Title, Count(Title) as 'Count' from combined group by Title order by count(Title)")

#All the rarer titles seem to be of rank or nobility, will combine these



combined <- sqldf(c("update combined set Title = 'Nobility' where Title in 
('Capt.',
'Col.',
'Don.',
'Dona.',
'Jonkheer.',
'Lady.',
'Major.',
'Mlle.',
'Mme.',
'Ms.',
'Sir.',
'the')","select * from combined"))

combined_age <- sqldf(c("update combined set Title = 'Nobility' where Title in 
('Capt.',
'Col.',
'Don.',
'Dona.',
'Jonkheer.',
'Lady.',
'Major.',
'Mlle.',
'Mme.',
'Ms.',
'Sir.',
'the')","select * from combined_age"))



```



#Seperate Train and Test for Variables we want

```{r}

train2 <- sqldf("select Pclass, Sex, SibSP, Parch, Title, AgeClass, Survived 
                from combined a 
                Where [Set] = 'train'")

test2 <- sqldf("select Pclass, Sex, SibSP, Parch, Title, AgeClass 
                from combined a 
                Where [Set] = 'test'")

train2_age <- sqldf("select Pclass, Sex, SibSP, Parch, Title, AgeClass, Survived 
                from combined_age a 
                Where [Set] = 'train'")

test2_age <- sqldf("select Pclass, Sex, SibSP, Parch, Title, AgeClass 
                from combined_age a 
                Where [Set] = 'test'")


```

#Model

```{r}

#Adding more cores to R/model for faster processing and leaving 1 for OS
cluster <- makeCluster(detectCores() - 1) 
registerDoParallel(cluster)

train2$Survived <- as.factor(train2$Survived)

train2_age$Survived <- as.factor(train2_age$Survived)

fitControl <- trainControl(method = "cv",
                           number = 2,
                           allowParallel = TRUE)



#Random Forest

#Comparing imputed age vs actual age

RFfitAge <- train(Survived ~., data = train2_age, method = "rf", trControl = fitControl)

RFfitAge$finalModel

print(RFfitAge)


RFfit <- train(Survived ~., data = train2, method = "rf", trControl = fitControl)

RFfit$finalModel

print(RFfit)


#Gradient Boosting Machine)

modFit <- train(Survived ~., method = "gbm", data=train2, verbose = FALSE)

modFit






#Predict the test set with the best model
#using modfit here

pred <- predict(RFfit, newdata = test2)

pred

submission <- as.data.frame(cbind(test$PassengerId, as.character(pred)))

names(submission) <- c("PassengerId","Survived")


write.csv(submission, file = "kagglescorerf.csv", row.names = FALSE)



```